---
title: "Analysis of Apple Mobility Data During the COVID-era"
author: "Nicole McHugh"
date: "2022-10-09"
output: html_document
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = TRUE)
```

## Setup

```{r set-parameters}
# set the state to work with
state_to_analyze <- "New Mexico"

# sanitizing the state name for output files
state_no_spaces <- gsub(state_to_analyze,
                        pattern = " ",
                        replacement = "_")

#set the name of the global data file to use
global_file_to_process <- "data/raw_data/applemobilitytrends-2022-04-12.csv"

# extract core name for output file naming use
core_name <- tools::file_path_sans_ext(basename(global_file_to_process))
```

```{r load-packages-functions}
# load the packages needed
library("dplyr")
library("readr")
library("ggplot2")
library("tidyr")
library("lubridate")
library("devtools")

# load the functions needed
source("code/functions/subset_data_to_a_state.R")
source("code/functions/counties_cities_count_function.R")
source("code/functions/wide_to_long_subset.R")
```

## Project Description

This rmarkdown fill use the `ggplot2` [@ggplot2] and `dplyr`[@dplyr] packages to analyze the Apple maps mobility data from the time before the COVID-19 pandemic to April 12, 2022.

The goal will be to create a subset to a particular state, tally the number of cities and counties in that state with mobility data, and to make several plots to visualize those data subsets.

The state that this document will analyze is `r state_to_analyze`.

## Subsetting data

```{r subset-state}
# subset out to the chosen state using the variables set above
state_subset <- subset_data_to_a_state(
  input_file_name = global_file_to_process,
  state_name = state_to_analyze)
```

## Wide to long data

```{r make-subset-long}
# convert the subsetted state data from wide to long format
long_subset <- wide_to_long_subset(
  input_file_name = paste0("output/subsetted_state_wide/",
                           state_no_spaces,
                           "_subset_",
                           core_name,
                           ".csv"))
```

## Tallied subset

```{r tallied-subset}
# tally up the number of cities and counties in the state that have mobility
# data available
tallied_subset <- counties_cities_count_function(
  input_file_name = paste0("output/subsetted_state_wide/",
                           state_no_spaces,
                           "_subset_",
                           core_name,
                           ".csv"))
```

## Figures
```{r plot-tallied}
# create a grouped barplot of the tallied data to include in the knitted
# documents, and save that figure to the output/figures directory a png
tallied_plot <- ggplot(data = tallied_subset,
                       aes(x = geo_type,
                           y = n,
                           fill = transportation_type)) +
  geom_col(position = position_dodge()) +
  labs(title = paste("Number of cities and/or counties in",
                     state_to_analyze,
                     "with mobility data"),
       x = "Type of area",
       y = "Number of areas")

ggsave(plot = tallied_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_tallied_plot.png"))

tallied_plot
```

```{r plot-time-series}
# produce a line plot of relative driving mobility across the state with date
# on the x axis and mean relative mobility on the y axis
timeseries_plot <- long_subset %>%
  filter(transportation_type == "driving") %>%
  group_by(date) %>%
  summarize(mean_mobility = mean(relative_mobility)) %>%
  ggplot(aes(x = lubridate::ymd(date),
             y = mean_mobility)) +
  geom_line() +
  labs(title = paste("Statewide mean relative mobility driving levels in",
                     state_to_analyze,
                     "during COVID"),
       x = "Date",
       y = "Mean relative mobility")

ggsave(plot = timeseries_plot,
       filename = paste0("output/figures/",
                         core_name,
                         "_",
                         state_no_spaces,
                         "_timeseries_plot.png"))
timeseries_plot

```

```{r scatter-plot}
# create scatter plot that maps all transportation types within the long
# subsetted data over time, save the figure as a png file in the output/figure
# folder, and display the figure in the knitted document
scatter_plot <- long_subset %>%
  group_by(transportation_type) %>%
  ggplot(aes(x = lubridate::ymd(date),
             y = relative_mobility,
            color = transportation_type)) +
  geom_point() +
  labs(title = paste(state_to_analyze,
                     "mobility scatter plot during COVID"),
       x = "Date",
       y = "Relative mobility")
ggsave(plot = scatter_plot,
       filename = paste0("output/figures/",
                         state_no_spaces,
                         core_name,
                         "_scatter_plot.png"))
scatter_plot

```


## Session Info

```{r session-info}
devtools::session_info()
```


## Sources Cited


